---
title: "Data Exploration"
author: "Madelyn Subali"
date: "5/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction 
The College Scorecard is a public-facing website that contains important information about colleges, including how much its graduates earn. Here, I will discuss on how the Scorecard shifted interest between high- and low-earnings colleges over the year. 

The College Scorecard was released at the start of September 2015. The report will show if colleges that predominantly grant Bachelor’s degrees, on the release of the Scorecard shift student's interest to high-earnings colleges relative to low-earnings ones. 

# Data Cleaning
```{r}
# Libraries
library(tidyverse)
library(fixest)
library(ggstance)
library(vtable)
library(multcomp)
library(purrr)
library(lubridate)
library(dplyr)
library(jtools)

#Reading in Google Trends Data
trends_files <- list.files(path = '.' , pattern='trends_', full.names=TRUE)
trends_files <- map_df(trends_files, read_csv) 

#Aggregating the Google Trends Data:
trends_files <- trends_files %>% 
   group_by(schname, keyword) %>%
  mutate(index_id = (index - mean(index, na.rm=TRUE)) /sd(index, na.rm=TRUE))

# Reading in Scorecard Data
score <- read_csv('Most+Recent+Cohorts+(Scorecard+Elements).csv')
id_name_link <- read_csv('id_name_link.csv')

# Remove university that share the same schname
id_name_link <- id_name_link %>%
  group_by(schname) %>% 
  mutate(n= n()) %>% 
  filter ( n == 1)

# Make column name to lower case to match with id_name_link file
names(score) <- tolower(names(score))

# Merging the scorecard data
join_files <- inner_join(id_name_link, trends_files, by='schname')
final_join <- inner_join(join_files, score, by='unitid')
write_csv(final_join, "clean_data.csv")
```

# Analysis 

According to a 2020 data retrieved from Northeastern University, the median weekly earnings of someone with a Bachelor's degree working in the United States is $1,248 with an annual earnings of $64,896. With this in mind, I will make $64,896 as a benchmark. Colleges with a median annual earning of $64,896 or higher 
will be marked as high-earning colleges and below $64,896 as low-earning colleges.

```{r} 
# Filter college for Bachelor's degree and removing NA values. 
clean_data <- read_csv('clean_data.csv')
bachelors_degree_trend <- clean_data %>% filter(preddeg == 3, na.rm=TRUE) 
```

Creating dummy variables with ifelse() function to set the benchmark, where '1' is high-earning and '0' is low-earning. 
```{r} 
bachelors_degree_trend$medianEarning <- ifelse(bachelors_degree_trend$`md_earn_wne_p10-reported-earnings` >= 64896, '1', '0') 
```

Selecting only the necessary columns that we want to use in our data set. 
```{r}
bachelors_degree_trend <- bachelors_degree_trend %>% 
                          dplyr::select(unitid, opeid.x, instnm, city, monthorweek, preddeg, sat_avg, ugds, `md_earn_wne_p10-reported-earnings`, medianEarning, index_id) %>%
                          mutate(date= as.Date(str_sub(monthorweek, 1,10))) %>%
                          group_by(month = floor_date(date, "month"), medianEarning) %>%
                          summarize(index_id = mean(index_id, na.rm = TRUE))
```

# Regression

According to the US department of education, the college scorecard was launched on September 12,2015. The regression below is going to test whether there is an impact on Google Trends index on median earnings, before and after the college scorecard was launched.

```{r} 
clean_data <- clean_data %>%
  mutate(earning = `md_earn_wne_p10-reported-earnings` >= 64896, post_scorecard = monthorweek >= 
           as.Date("2015-09-12"))

m1 <- feols(index_id ~ earning*post_scorecard, data=clean_data) 
export_summs(m1, digits = 3)
```

### Interpretation of Regression: 
From the regression above, we can determine that there is an 1.9 percent change in the effect of high-earning colleges in the trends after the scorecard was released. 

# Visualization
The visualization below will show the impacts of Google Trends index within the years of high-earning and low-earning colleges. Here, high-earning is depicted as '1', while low-earning is depicted as '0'.

```{r, echo=FALSE}
ggplot(bachelors_degree_trend, aes(x=month, y=index_id, color=medianEarning)) + 
  geom_line(aes(color= medianEarning), size=0.8) +
  geom_vline(xintercept = as.Date("2015-09-12")) + theme_bw() + guides(color=guide_legend("Median Earning"))+
  labs(title= "The Impact of Google Trends Index Within the Years for Colleges")
```

### Interpretation for Visualization: 
From the line graph, we can conclude that there is not much difference between the low-earning and high-earning. We can see that both lines follow a similar pattern where it decreases over time, overall. In the date that the college scorecard was launched, which is 12 September 2015, we can see that both high-earning and low-earning colleges had an increase in a short period of time before decreasing over the next year. This suggest that median earnings for colleges is not significantly impacted by the Google trends index.
